Informative Oceanographic Sampling using Excursion Probabilities for Multivariate Random Fields

### 2019-09-13 main_new.tex
updates

The solution to the second part of the equation above, is then to
evaluate a four-variable cumulative distribution function associated
with Eqn. (\ref{part1:phi4}), accounting for the mean, variance and
covariances of $\bZ_1$, $\bZ_2$ and $\bm$ in the appropriate linear
combinations. Using vector-matrix formulations, we define linear
combinations:
\begin{equation}\label{Vcomb}
    \bV=\left[
    \begin{array}{cccccc}
        s_a & 0 & 0 & 0 & 1 & 0 \\
         0 & s_b & 0 & 0 & 0 & 1 \\
         0 & 0 & s_a & 0 & 1 & 0\\
         0 & 0 & 0 & s_b & 0 & 1\\
    \end{array}
    \right] 
    \left(
    \begin{array}{cccccc}
         Z_{1,a} \\
         Z_{1,b} \\
         Z_{2,a} \\
         Z_{2,b} \\
         m_{a} \\
         m_{b}
    \end{array}
    \right),
\end{equation}
and this part of the equation then becomes:
\begin{eqnarray}\label{part1:phi4}
&& P \left( Z_{1,a} \leq \frac{t_a-m_{a}}{s_{a}}, 
Z_{1,b} \leq \frac{t_b-m_{b}}{s_{b}},Z_{2,b} \leq \frac{t_a-m_{a}}{s_{a}}, 
Z_{2,b} \leq \frac{t_b-m_{b}}{s_{b}} \right)  \\
&=&  \Phi_4 
\begin{pmatrix} 
\begin{bmatrix} t_a\\
t_b \\
t_a \\
t_b
\end{bmatrix};
E( \bV),Var(\bV)
\end{pmatrix} \nonumber.
\end{eqnarray}

The multivariate cumulative probabilities in two and four dimensions
are fast to compute using methods of \cite{genz2009computation}.  In
the end, the first part in Eq. (\ref{two_parts0}) and the second in
Eq. (\ref{part1:phi4}) are computed for each location on the grid, and
summing over the $x \in \mathcal{M}_g$, gives the design evaluation
criterion in Eq. (\ref{sur}). \kc{This might be a prevalent style in
  Math/Statistics, but this entire formulation is way too dense. I
  think it might be a good idea to have some basic intuition worked
  out early before the start of Sec 3}

@@

In our application with temperature and salinity, the solutions to the complex integral equations for the expected IBV are reduced to calculating bi-variate and four-variate cumulative distribution functions for Gaussian vectors. For the more general situation one would be interested in $K$ random fields. Say, in the ocean sciences the variables could include chlorophyll or other bio-chemical quantities, in addition to temperature and salinity.

The equations derived in the previous section can then be generalized. Let $\bxi(\bx)=(\xi_1(\bx),\ldots,\xi_K(\bx))$, and the expected Bernoulli variance is
\begin{equation}\label{two_partsK}
E_{\by}(p(\bx;\by)(1-p(\bx;\by))), \hspace{3mm} p(\bx;\by)=P(\xi_k(\bx) \leq t_k ; k=1,\ldots,K|\by). 
\end{equation}
By the first key result of linear combination of Gaussian variables, we reduce the integral to a $K$ dimensional integral of the relevant $\bm=E(\bxi(\bx)|\by)$. Next, reducing the integral to two parts, we standardize the variables and compute cumulative distributions. 
In summary, we get 
\begin{equation}\label{two_parts}
E_{\by}(p(\bx)(1-p(\bx))) =  \Phi_K 
\begin{pmatrix}
\begin{bmatrix} t_1\\
\vdots \\
t_K 
\end{bmatrix};
\bmu,\bS+\bPsi 
\end{pmatrix}
- \Phi_{2K} 
\begin{pmatrix}
\begin{bmatrix} t_1\\
\vdots \\
t_K \\
t_1\\
\vdots \\
t_K 
\end{bmatrix};
E(\bV),Var(\bV) 
\end{pmatrix},\nonumber
\end{equation}
where $\bS$ and $\bPsi$ are direct $K$-variate extensions of the matrices in Eq. (\ref{two_parts0}), and the matrix $\bV$ finds the relevant linear combination of two independent vectors $\bZ_l=(Z_{l,1},\ldots,Z_{l,K})$, $l=1,2$ and $\bm$, as an extension of Eq. (\ref{Vcomb}).

>>>>

The solution to the second part of the equation above, is then to
evaluate a four-variable cumulative distribution function associated
with Eqn. (\ref{part1:phi4}), accounting for the mean, variance and
covariances of $\bZ_1$, $\bZ_2$ and $\bm$ in the appropriate linear
combinations. Using vector-matrix formulations, we define linear
combinations:
\begin{equation}\label{Vcomb}
    \bV=\left[
    \begin{array}{cccccc}
        s_a & 0 & 0 & 0 & 1 & 0 \\
         0 & s_b & 0 & 0 & 0 & 1 \\
         0 & 0 & s_a & 0 & 1 & 0\\
         0 & 0 & 0 & s_b & 0 & 1\\
    \end{array}
    \right] 
    \left(
    \begin{array}{cccccc}
         Z_{1,a} \\
         Z_{1,b} \\
         Z_{2,a} \\
         Z_{2,b} \\
         m_{a} \\
         m_{b}
    \end{array}
    \right),
\end{equation}
and this in turn then becomes:
\begin{eqnarray}\label{part1:phi4}
&& P \left( Z_{1,a} \leq \frac{t_a-m_{a}}{s_{a}}, 
Z_{1,b} \leq \frac{t_b-m_{b}}{s_{b}},Z_{2,b} \leq \frac{t_a-m_{a}}{s_{a}}, 
Z_{2,b} \leq \frac{t_b-m_{b}}{s_{b}} \right)  \\
&=&  \Phi_4 
\begin{pmatrix} 
\begin{bmatrix} t_a\\
t_b \\
t_a \\
t_b
\end{bmatrix};
E( \bV),Var(\bV)
\end{pmatrix} \nonumber.
\end{eqnarray}

The multivariate cumulative probabilities in two and four dimensions
are fast to compute using methods of \cite{genz2009computation}.  In
the end, the first part in Eq. (\ref{two_parts0}) and the second in
Eq. (\ref{part1:phi4}) are computed for each location on the grid, and
summed over $x \in \mathcal{M}_g$, gives the design evaluation
criterion in Eq. (\ref{sur}). \kc{This might be a prevalent style in
  Math/Statistics, but this entire formulation is way too dense. I
  think it might be a good idea to have some basic intuition worked
  out early before the start of Sec 3.}

@@

In our application domain with the variables temperature and salinity,
the solutions to the complex integral equations for the expected IBV
are reduced to calculating bi-variate and four-variate cumulative
distribution functions for Gaussian vectors. For the more general
situation one would be interested in $K$ random fields. Say, in the
ocean sciences the variables could include chlorophyll or other
bio-geochemical quantities, in addition to temperature and salinity.

The equations derived in the previous section can then be generalized
\kc{Aha! Looks like this is where the generalization across multiple
  variables can be brought out. If so, state this up front that we're
  starting off with 2 and then generalizing to more variables.}. Let
$\bxi(\bx)=(\xi_1(\bx),\ldots,\xi_K(\bx))$, and the expected Bernoulli
variance is:
\begin{equation}\label{two_partsK}
E_{\by}(p(\bx;\by)(1-p(\bx;\by))), \hspace{3mm} p(\bx;\by)=P(\xi_k(\bx) \leq t_k ; k=1,\ldots,K|\by). 
\end{equation}
By the first key result of linear combination of Gaussian variables,
we reduce the integral to a $K$ dimensional integral of the relevant
$\bm=E(\bxi(\bx)|\by)$. Next, reducing the integral to two parts, we
standardize the variables and compute cumulative distributions.  In
summary, we get:
\begin{equation}\label{two_parts}
E_{\by}(p(\bx)(1-p(\bx))) =  \Phi_K 
\begin{pmatrix}
\begin{bmatrix} t_1\\
\vdots \\
t_K 
\end{bmatrix};
\bmu,\bS+\bPsi 
\end{pmatrix}
- \Phi_{2K} 
\begin{pmatrix}
\begin{bmatrix} t_1\\
\vdots \\
t_K \\
t_1\\
\vdots \\
t_K 
\end{bmatrix};
E(\bV),Var(\bV) 
\end{pmatrix},\nonumber
\end{equation}
where $\bS$ and $\bPsi$ are direct $K$-variate extensions of the matrices in Eq. (\ref{two_parts0}), and the matrix $\bV$ finds the relevant linear combination of two independent vectors $\bZ_l=(Z_{l,1},\ldots,Z_{l,K})$, $l=1,2$ and $\bm$, as an extension of Eq. (\ref{Vcomb}).
